Human Judgement Forecasting of COVID-19 in the UK

### 2023-02-28 manuscript/manuscript.tex
second test commit to test overleaf sync
comments from Sam on manuscript

\begin{abstract}

%Abstracts should be up to 300 words and provide a succinct summary of the article. Although the abstract should explain why the article might be interesting, care should be taken not to inappropriately over-emphasise the importance of the work described in the article. Citations should not be used in the abstract, and the use of abbreviations should be minimized. If you are writing a Research or Systematic Review article, please structure your abstract into Background, Methods, Results, and Conclusions.

@@

Accurate forecasts of infectious diseases can aid public health decision-making. In the past, two studies found ensembles of human judgement forecasts of COVID-19 to show predictive performance comparable to ensembles of computational models, at least when predicting case incidences. We present a follow-up to a study conducted in Germany and Poland and investigate a novel approach to combine human judgement and epidemiological modelling.

@@

From May 24th to August 16th 2021, we elicited weekly one to four-week-ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. Forecasts were elicited as part of a tournament, the "UK Crowd Forecasting Challenge", and a median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$. This forecast was automatically mapped to cases and deaths using the renewal equation and a convolution over a delay distribution. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

@@

The ensemble of human forecasters overall performed comparably to the offical European Forecast Hub ensemble on both cases and deaths, although results were susceptible to changes in details of the evaluation. $R_t$ foreasts performed comparably to direct forecasts ona cases, but worse on deaths. Self-identified experts did not outperform non-experts.

@@

Human judgement can be useful in predicting the large-scale evolution of an infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a "good" forecast is dependent on the forecast consumer.

@@

Infectious disease modelling and forecasting has attracted wide-spread attention during the COVID-19 pandemic and helped inform decision making in public health organisations and governments \cite{cramerEvaluationIndividualEnsemble2021, venkatramananUtilityHumanJudgment2022}. 
% Beginning in March 2020, forecasts for different COVID-19 targets have been systematically collated by Forecast Hubs in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. 
Most forecasts were based on computational models of COVID-19, but some authors also explored human judgement forecasting as an alternative \cite{recchiaHowWellDid2021, mcandrewExpertJudgmentModel2022, bosseComparingHumanModelbased2022, mcandrewChimericForecastingCombining2022}.

Past research found that in the context of infectious disease forecasting, human judgement forecasts could achieve predictive performance broadly comparable to forecasts generated based on mathematical modelling, in particular when forecasting incident cases. \citet{farrowHumanJudgmentApproach2017} found that an aggregate of human predictions outperformed computational models when predicting the 2014/15 and 2015/16 flu season in the US. However, a comparable approach performed slightly worse than computational models at predicting the 2014/15 outbreak of chikungunya in the Americas. 
\citet{bosseComparingHumanModelbased2022} found an ensemble of human forecasters to outperform an ensemble of computational models when predicting cases of COVID-19 in Germany, but performing worse when predicting incident deaths. Similarly, \citet{mcandrewChimericForecastingCombining2022} reported an ensemble of human forecasters to perform comparably to an ensemble of computational models when predicting incident cases, and worse when predicting incident deaths. \citet{farrowHumanJudgmentApproach2017} and in particular \citet{bosseComparingHumanModelbased2022} struggled to recruit a large number of participants (numbers of active forecasters ranged from 22 to 61 in \citet{mcandrewChimericForecastingCombining2022}, 7 to 24 in \citet{farrowHumanJudgmentApproach2017}, and 4 to 10 in \citet{bosseComparingHumanModelbased2022}).

In some situations, human judgement forecasting has clear advantages relative to computational models. Humans can very quickly provide forecasts, even in situations where data is sparse and lots of parameters are unknown. In addition, humans are able to answer a broad set of question (such as for example the likelihood that a given actor will take some specified action) or can take factors into account that are hard to encode in a computational model. On the other hand, human judgement forecasting is difficult to scale due to the time and effort required, and humans may be at a disadvantage at tasks that require complex computations.

@@

Methods that aim to combine human judgement and mathematical modelling are therefore appealing. One straightforward way is to combine separate human judgement and computational model forecasts with the goal of improving predictive performance \citep{mcandrewChimericForecastingCombining2022}. \cite{farrowHumanJudgmentApproach2017, bosseComparingHumanModelbased2022} and others suggested additional possibilities in the context of infectious diseases that may also help reduce the amount of human effort required. One approach is to use human forecasts, for example of relevant disease parameters, as an input to computational modelling. Another approach is to use mathematical modelling as an input to human judgement, for example by giving experts the option to make post-hoc adjustments to model outputs. \citet{bosseComparingHumanModelbased2022} proposed to ask human forecasters to predict the effective reproduction number $R_t$ (the average number of people an infected person would infect in turn) and to feed this forecast into a mathematical model in order to obtain forecasts for case and death numbers.

This paper represents a follow-up study to \citet{bosseComparingHumanModelbased2022} in the United Kingdom with forecasts made over the course of thirteen weeks between May and August 2021. Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a mathematical model.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022a} elicits predictions for various COVID-19 related forecast targets from different research groups every week. Forecasts had to be submitted every Monday 23.59pm GMT. Forecasts are made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epiweeks, ending on a Saturday and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub follow a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecasts submitted to the hub were automatically checked for conformity with the required format and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble shown on all official Forecast Hub visualisiations (https://covid19forecasthub.eu/) was a mean-ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July on, the default Forecast Hub ensemble became a median-ensemble. The number of models included in the Forecast Hub ensemble is shown in Figure \ref{fig:num-forecasters} in the SI.

@@

The web application offered participants two different ways of making a forecast, called 'direct' (or 'classical') and 'Rt' forecast. To make a 'direct' forecast (as described in more detail in \cite{bosseComparingHumanModelbased2022}), participants selected a predictive distribution (by default a log-normal distribution) and adjusted the median and width of the distribution to change the central estimate and uncertainty. In addition to information about past observations, participants could see various metrics and data sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

\textbf{SAM TO THE RESCUE}

In addition to the 'direct' forecasts, we implemented a second forecasting method ('Rt forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$. This forecast was then translated into forecasts of cases using the renewal equation \citep{fraserEstimatingIndividualHousehold2007} as implemented in the \textsf{R} package \texttt{EpiNow2} \citep{epinow2} (see details in the SI). In order to obtain forecasts for deaths, we convolved predicted infections as implied by the $R_t$ forecast over data-based delay distributions \cite{epinow2, sherrattExploringSurveillanceData2021, abbottEstimatingTimevaryingReproduction2020a} to model the time between infection and report date. As $R_t$-estimates up to two weeks prior to the forecast data were deemed uncertain, we also asked participants to submit an estimate of $R_t$ for the two weeks prior to the current forecast date. Participants were therefore asked to estimate/predict six $R_t$ values, four of them in the future. Upon pressing a button, participants could see a preview of the evolution of cases implied by their current $R_t$ forecast. However, due to the high computational complexity and necessary waiting times, participants could not preview the death forecast implied by their current input for $R_t$. Figure \ref{fig:screenshot-rt} shows a screenshot of the forecast interface for $R_t$ forecasts.

@@

\subsection*{EpiNow2?}

SAM TO THE RESCUE, RELEVANT PART OF OLD PAPER PASTED HERE AND COMMENTED OUT. DO WE NEED AN EXTRA SECTION, OR SHOULD THIS BE INCLDUED ABOVE?

% From old paper: 
% The first of these models, here called “renewal model”, used the renewal equation [41] to predict reported cases and deaths (see details in S1 Text). It estimated the effective reproduction number Rt (the average number of people each person infected at time t is expected to infect in turn) and modelled future infections as a weighted sum of past infection multiplied by Rt. Rt was assumed to stay constant beyond the forecast date, roughly corresponding to continuing the latest exponential trend in infections. On the 9th of November we altered the date when Rt was assumed to be constant from two weeks prior to the date of the forecast to the forecast date, which we found to yield a more stable Rt estimate. Reported case and death notifications were obtained by convolving predicted infections over data-based delay distributions [40, 42–44] to model the time between infection and report date. The renewal model was used to predict cases as well as deaths with forecasts being generated for each target separately. Death forecasts from the renewal model were therefore not informed by past cases. One submission of the renewal model on December 28th 2020 was delayed and therefore not included in the official Forecast hub ensemble.
% 
% Line list data used to inform the prior for the delay from symptom onset to test positive case report or death in the model-based forecasts was sourced from [45] with data available up to the 1st of August. All model fitting was done using Markov-chain Monte Carlo (MCMC) in stan [46] with each location and forecast target being fitted separately.

@@

In order to boost participation compared to our last crowd forecasting study in Germany and Poland \citep{bosseComparingHumanModelbased2022}, we announced an official tournament, the "UK Crowd Forecasting Challenge". Participants were asked to submit weekly predictions for reported cases and deaths linked to COVID-19 in the United Kingdom one to four weeks into the future. Everyone who had submitted a forecast for targets in the UK during the tournament period from the 24th of May 2021 to the 16th of August 2021 was deemed a participant and eligible for a prize. The first prize was 100 GBP, second prize 50 GBP and third prize 25 GBP. Participant performance was determined using the mean weighted interval score (WIS) on the log scale (see details in the next Section, %\ref{sec:analysis}), 
averaged across forecast dates, horizons and forecast targets. For the purpose of the tournament ranking, participants who did not submit a forecast in a given week were assigned the median score of all other participants who submitted a forecast that week. The UK crowd forecasting challenge was announced over Twitter and our networks. 
% (e.g. https://www.voice-global.org/public/opportunities/archived/uk-covid-19-crowd-forecasting-challenge/)
In addition, we created a project website, \url{https://crowdforecastr.org}, made weekly posts on Twitter and sent participants who had registered on the online application weekly emails with a reminder and a summary of their past performance. A public leaderboard was available on our website \url{https://epiforecasts.io}. Participants could choose to make a direct forecast as well as an $R_t$ forecast and were counted as two separate forecasters and eligible for prizes twice. Weekly forecasts had to be submitted between Sunday 12pm and Monday 8pm UK time.

@@

Our analysis of the data largely follows the one outlined in \citet{bosseComparingHumanModelbased2022}. If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs, which found predictive performance to be poor and unreliable beyond this horizon \cite{cramerEvaluationIndividualEnsemble2021, sherrattPredictivePerformanceMultimodel2022a, bracherShorttermForecastingCOVID192021}. We analysed all forecasts stratified by forecast target (cases or deaths), forecast horizon, and forecast approach. We compared the performance of the direct vs. $R_t$ forecasting approach using instances where we had both a direct forecasts and an $R_t$ forecast from the same person.

@@

The median number of unique participants in any given week was 17, the minimum was 6 and the maximum was 51 (compared to 6, 2, and 10 in \cite{bosseComparingHumanModelbased2022}). With respect to the number of submissions from an individual participant, we observed similar patterns as \cite{bosseComparingHumanModelbased2022}: An individual forecaster participated on average in 2.6 weeks out of 13. The median number of submissions from a single individual was one, meaning that most forecasters dropped out after their first submission. Only five participants submitted a forecast in ten or more weeks and only two submitted a forecast in all thirteen weeks, one of whom is an author on this study. The number of direct forecasts was higher than the number of $R_t$ forecasts in all weeks (see Figure \ref{fig:num-forecasters}).

@@

\begin{figure*}
\centering
\includegraphics[width=0.99\textwidth]{../output/figures/performance.png}
\caption{\bf{Predictive performance across forecast horizons.} A-D: WIS stratified by forecast horizon for cases and deaths on the natural and log scale. E, F: Empirical coverage of the 50\% and 90\% prediction intervals stratified by forecast horizon and target type.}
\label{fig:performance}
\end{figure*}

@@

In terms of the WIS on the natural scale, direct human forecasts performed best at a two-week forecast horizon, while $R_t$ forecasts performed significantly worse than all other approaches (see Figure \ref{fig:performance} and Table \ref{tab:scores}). From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August, $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). For four-week-ahead death forecasts, the Hub ensemble performed slightly better than other approaches. 
On the natural scale, the highest scores were concentrated towards the end of the study period, where death incidences were highest, while on the log scale, scores were more evenly distributed across the study period. On the log scale, direct human forecasts, the combined crowd ensemble and the Forecast Hub ensemble showed comparable performance, while $R_t$ forecasts performed again significantly worse. 
%Note that the picture changes when looking at the revised data which was corrected upwards compared to the original data(DO WE WANT TO DO THAT? IS THAT MEANINGFUL?) (see Figure \ref{fig:data-revisions}). For the revised data, direct forecasts performed significantly worse than both the Forecast Hub ensemble and the $R_t$ forecasts on the log scale, while still performing better than the $R_t$ forecasts on the natural scale (see Table \ref{tab:scores-revised}). 
%For revised data: On the natural scale, the largest scores occurred at the end of the study period when deaths were highest. On the log scale, however, underprediction in June and July when death incidences were lower, received more weight and direct crowd forecast outperformed $R_t$ forecasts. 
$R_t$ forecasts show strong underdispersion with empirical coverage of the 50\% prediction intervals being close to zero and smaller than 50\% for the 90\% prediction intervals (see Figure \ref{fig:performance} and Table \ref{tab:scores}). The Forecast Hub ensemble was overdispersed with empirical coverage exceeding nominal coverage, and for the direct human forecasts, empirical coverage was close to nominal coverage.

@@

For cases, where participants could observe the case forecast implied by their $R_t$ forecast, predictive performance was similar between corresponding direct and $R_t$ forecasts for most forecasters who had submitted both (see Figure \ref{fig:comparison-direct-rt-individual}). For deaths, predictive performance of their $R_t$ forecasts was significantly worse than their corresponding direct forecasts for most forecasters, suggesting that the convolution from latent infections implied by the $R_t$ forecasts to deaths did not capture the relationship between cases and deaths accurately. This inadequacy may be related to the change in the case fatality ratio over the course of the study period due to the rise of the Delta variant starting in May 2021 in the UK.  
%IMPORTANTLY THOUGH, THE PICTURE DOES LOOK DIFFERNT WHEN LOOKING AT THE REVISED DATA.

% When looking at an ensemble including only those data points where a single forecaster has submitted both a direct and an $R_t$ forecast, we found a similar picture. Performance for cases was broadly comparable, with direct forecasts slightly better than $R_t$ forecasts on the natural scale and slightly worse on the log scale. On the log scale, the ensemble of $R_t$ forecasts performed significantly worse than the ensemble of corresponding direct forecasts both on the natural and on the log scale.

@@

In this paper, we presented a follow-up study to \cite{bosseComparingHumanModelbased2022}, analysing human judgement forecasts of cases of and deaths from COVID-19 in the United Kingdom submitted to the European COVID-19 Forecast Hub between the 24th of May and the 16th of August 2021. Human judgement forecasts were generated using two different forecasting approaches, a) direct forecasts of cases and deaths and b) forecasts of the effective reproduction number $R_t$, which were mapped to cases using the renewal equation and convolved over a delay distribution to obtain deaths. We found that human judgement forecasts of case numbers, regardless of how they were obtained, showed performance broadly comparable to the ensemble of all forecasts submitted to the European Forecast Hub. For deaths, performance of direct human forecasts was comparable to that of the Forecast Hub ensemble, while $R_t$ forecasts performed significantly worse. Model rankings were not always robust, but sometimes changed depending on how forecasts are evaluated. For example, model rankings could differ when focusing on a different forecast horizon, or analysing WIS values on the natural scale vs. on the log scale (i.e. after log-transforming both forecasts and observations).

>>>>

\begin{abstract}
%Abstracts should be up to 300 words and provide a succinct summary of the article. Although the abstract should explain why the article might be interesting, care should be taken not to inappropriately over-emphasise the importance of the work described in the article. Citations should not be used in the abstract, and the use of abbreviations should be minimized. If you are writing a Research or Systematic Review article, please structure your abstract into Background, Methods, Results, and Conclusions.

@@

Accurate forecasts of infectious diseases can aid public health decision-making. In the past, two studies found ensembles of human judgement forecasts of COVID-19 to show predictive performance comparable to ensembles of computational models, at least when predicting case incidences. We present a follow-up to a study conducted in Germany and Poland and investigate a novel joint approach to combine human judgement and epidemiological modelling.

@@

From May 24th to August 16th 2021, we elicited weekly one to four-week-ahead forecasts of cases and deaths from COVID-19 in the UK from a crowd of human forecasters. Forecasts were elicited as part of a tournament, the "UK Crowd Forecasting Challenge", and a median ensemble of all forecasts was submitted to the European Forecast Hub. Participants could use two distinct interfaces: in one, forecasters submitted a predictive distribution directly, in the other forecasters instead submitted a forecast of the effective reproduction number $R_t$. This forecast was automatically mapped to cases and deaths using the \texttt{EpiNow2} \textsf{R} package and associated model. Forecasts were scored using the weighted interval score on the original forecasts, as well as after applying the natural logarithm to both forecasts and observations.

@@

The ensemble of human forecasters overall performed comparably to the official European Forecast Hub ensemble on both cases and deaths, although results were susceptible to changes in details of the evaluation. $R_t$ forecasts performed comparably to direct forecasts on cases, but worse on deaths. Self-identified experts did not outperform non-experts though the sample size was small.

@@

Human judgement can be useful in predicting the large-scale evolution of an infectious disease such as COVID-19. The results of forecast evaluations can change depending on what metrics are chosen and judgement on what does or doesn't constitute a "good" forecast is dependent on the forecast consumer. Combinations of human and computational forecasts hold potential but present real-world challenges that need to be solved.

@@

Infectious disease modelling and forecasting has attracted wide-spread attention during the COVID-19 pandemic and helped inform decision making in public health organisations and governments \cite{cramerEvaluationIndividualEnsemble2021, venkatramananUtilityHumanJudgment2022}. 
% Beginning in March 2020, forecasts for different COVID-19 targets have been systematically collated by Forecast Hubs in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. 
Most forecasts used to inform decision making were based on computational models of COVID-19, but some authors also explored human judgement forecasting as an alternative or in combination \cite{recchiaHowWellDid2021, mcandrewExpertJudgmentModel2022, bosseComparingHumanModelbased2022, mcandrewChimericForecastingCombining2022}.

Past research found that in the context of infectious disease forecasting, human judgement forecasts could achieve predictive performance broadly comparable to forecasts generated based on mathematical modelling, in particular when forecasting incident cases. \citet{farrowHumanJudgmentApproach2017} found that an aggregate of human predictions outperformed computational models when predicting the 2014/15 and 2015/16 flu season in the US. However, a comparable approach performed slightly worse than computational models at predicting the 2014/15 outbreak of chikungunya in the Americas. 
\citet{bosseComparingHumanModelbased2022} found an ensemble of human forecasters to outperform an ensemble of computational models when predicting cases of COVID-19 in Germany, but performing worse when predicting incident deaths. Similarly, \citet{mcandrewChimericForecastingCombining2022} reported an ensemble of human forecasters to perform comparably to an ensemble of computational models when predicting incident cases, and worse when predicting incident deaths. \citet{farrowHumanJudgmentApproach2017} and in particular \citet{bosseComparingHumanModelbased2022} struggled to recruit a large number of participants (numbers of active forecasters ranged from 22 to 61 in \citet{mcandrewChimericForecastingCombining2022}, 7 to 24 in \citet{farrowHumanJudgmentApproach2017}, and 4 to 10 in \citet{bosseComparingHumanModelbased2022}). 
It is also important to note that in the majority of these studies it is difficult to unpick the reliance of human forecasters on the computational models they were being compared to or to similar models so we should expect performance to at least equal with computational models in most settings.

In some situations, human judgement forecasting has clear advantages relative to computational models. Humans can very quickly provide forecasts, even in situations where data is sparse and lots of parameters are unknown. In addition, humans are able to answer a broad set of question (such as for example the likelihood that a given actor will take some specified action) or can take factors into account that are hard to encode in a computational model. On the other hand, human judgement forecasting is difficult to scale due to the time and effort required, humans may be at a disadvantage at tasks that require complex computations. Also, the use of human judgement forecasts by decision makers may be complicated by the lack of understanding of the basis on which they were made, if forecasters don't also provide a rational for their predictions.

@@

Methods that aim to combine human judgement and mathematical modelling are therefore appealing, though we note that presenting mathematical modelling as truly independent is perhaps misleading as the vast majority of computational models in use in epidemiology have at least some element of human judgement supporting their structure or usage. One straightforward method is to combine separate human judgement and computational model forecasts with the goal of improving predictive performance in the form of an ensemble. This has been widely shown to improve performance across model types \citep{mcandrewChimericForecastingCombining2022}. \cite{farrowHumanJudgmentApproach2017, bosseComparingHumanModelbased2022} and others suggested additional possibilities in the context of infectious diseases that may also help reduce the amount of human effort required. One approach is to use human forecasts, for example of relevant disease parameters, as an input to computational modelling. Another approach is to use mathematical modelling as an input to human judgement, for example by giving experts the option to make post-hoc adjustments to model outputs. \citet{bosseComparingHumanModelbased2022} proposed to ask human forecasters to predict the effective reproduction number $R_t$ (the average number of people an infected person would infect in turn) and to then feed this forecast into a mathematical simulation model in order to obtain forecasts for case and death numbers.

This paper represents a follow-up study to \citet{bosseComparingHumanModelbased2022} in the United Kingdom with forecasts made over the course of thirteen weeks between May and August 2021. Forecasts were elicited from experts and laypeople as part of a public forecasting tournament, the "UK Crowd Forecasting Challenge" using a web application. All forecasts were submitted to the European COVID-19 Forecast Hub, one of several Forecast Hubs that have been systematically collating forecasts of different COVID-19 forecast targets in the US \citep{cramerEvaluationIndividualEnsemble2021}, Germany and Poland \citep{bracherShorttermForecastingCOVID192021, bracherNationalSubnationalShortterm2021}, and Europe \citep{sherrattPredictivePerformanceMultimodel2022a}. This study aims to investigate whether the original findings in \citet{bosseComparingHumanModelbased2022} with respect to forecaster performance replicate in a different country and with an increased number of participants. In addition, it explores the approach proposed in \citet{bosseComparingHumanModelbased2022} to ask participants for a forecast of the effective reproduction number $R_t$ which is then translated into a forecast of cases and deaths using a mathematical model. We describe this approach as human in the loop computational modelling and consider it a formalisation of often practiced manual intervention in computational forecasts.

@@

The European COVID-19 Forecast Hub \cite{sherrattPredictivePerformanceMultimodel2022a} elicits predictions for various COVID-19 related forecast targets from different research groups every week. Forecasts had to be submitted every Monday 23.59pm GMT. Forecasts are made for incident weekly reported numbers of cases of and deaths from COVID-19 on a national level for various European countries over a one to four week forecast horizon. While forecasts were submitted on Mondays, weeks were defined as epiweeks, ending on a Saturday and starting on Sunday. Forecast horizons were therefore in fact 5, 12, 19 and 26 days. Submissions to the European Forecast Hub follow a quantile-based format with 23 quantiles of each output measure at levels 0.01, 0.025, 0.05, 0.10, 0.15,…, 0.95, 0.975, 0.99.
Every week, forecasts submitted to the hub were automatically checked for conformity with the required format and all eligible forecasts combined into different ensembles. Until the 12th of July 2021 the default Hub ensemble shown on all official Forecast Hub visualisations (https://covid19forecasthub.eu/) was a mean-ensemble (i.e., the $\alpha$-quantile of the ensemble is given by the mean of all submitted $\alpha$-quantiles). From the 29th of July on, the default Forecast Hub ensemble became a median-ensemble. The median number of models included in the Forecast Hub ensemble was nine for cases and ten for deaths (see Figure \ref{fig:num-forecasters} in the SI).

@@

The web application offered participants two different ways of making a forecast, called 'direct' (or 'classical') and 'Rt' forecast. To make a 'direct' forecast (as described in more detail in \cite{bosseComparingHumanModelbased2022}), participants selected a predictive distribution (by default a log-normal distribution) and adjusted the median and width of the distribution to change the central estimate and uncertainty. In addition to information about past observations, participants could see various metrics and data such as the test positivity rate and vaccination rate sourced from Our World in Data \citep{owidcoronavirus}. Figure \ref{fig:screenshot-classical} shows a screenshot of the forecast interface for direct forecasts.

In addition to the 'direct' forecasts, we implemented a second forecasting method ('Rt forecasts'), where we asked participants to make a forecast of the effective reproduction number $R_t$ based on a baseline forecast produced by the \texttt{EpiNow2} \textsf{R} package effective reproduction number model \citep{epinow2} which we also used in \cite{bosseComparingHumanModelbased2022} as a standalone computational model. This forecast was then translated into forecasts of cases using the simulation model from the \texttt{EpiNow2} \textsf{R} package \citep{epinow2}. \texttt{EpiNow2} implements a renewal equation based \citep{fraserEstimatingIndividualHousehold2007}  generative process for latent infections where the effective reproduction number is modelled as a Gaussian process. These latent infections are then convolved with delay distributions, such as the incubation period and reporting delay, and assumed to follow a negative binomial observation model with a day of the week effect to produce an estimate of reported cases. This approach has been widely used for short-term forecasting \cite{bosseComparingHumanModelbased2022} [CITE ECDC] [CITE SPI-M paper] and used to inform policy makers via reproduction number estimates [CITE one of the SPI-M reproduction number summary papers or kaths paper on multiple surveillance Rt esimates]. Further details are given in the SI.

In order to obtain forecasts for deaths, we similarly convolved observed and predicted cases as implied by the $R_t$ forecast over data-based delay distributions \cite{sherrattExploringSurveillanceData2021, abbottEstimatingTimevaryingReproduction2020a} to model the time between infection and report date  and applied a case fatality ratio estimated based on the relationship between observed and estimated reported cases and deaths again using the \texttt{EpiNow2} \textsf{R} package \citep{epinow2}. Again further details are given in the SI and in our previous work \citetp{bosseComparingHumanModelbased2022}.

As $R_t$-estimates up to two weeks prior to the forecast data were uncertain due to the delay between infection and case report, we also asked participants to submit an estimate of $R_t$ for the two weeks prior to the current forecast date. Participants were therefore asked to estimate/predict six $R_t$ values, four of them beyond the forecast horizon. Upon pressing a button, participants could see a preview of the evolution of cases implied by their current $R_t$ forecast. However, due to the computational complexity and resource limitations, participants could not preview the death forecast implied by their current input for $R_t$ nor could they influence the assumed case fatality ratio or delay between reported cases and reported deaths. Figure \ref{fig:screenshot-rt} shows a screenshot of the forecast interface for $R_t$ forecasts.

@@

@@

In order to boost participation compared to our last crowd forecasting study in Germany and Poland \citep{bosseComparingHumanModelbased2022} which struggled in this regard, we announced an official tournament, the "UK Crowd Forecasting Challenge". Participants were asked to submit weekly predictions for reported cases and deaths linked to COVID-19 in the United Kingdom one to four weeks into the future. Everyone who had submitted a forecast for targets in the UK during the tournament period from the 24th of May 2021 to the 16th of August 2021 was deemed a participant and eligible for a prize. The first prize was 100 GBP, second prize 50 GBP and third prize 25 GBP. Participant performance was determined using the mean weighted interval score (WIS) on the log scale (see details in the next Section, %\ref{sec:analysis}), 
averaged across forecast dates, horizons and forecast targets. For the purpose of the tournament ranking, participants who did not submit a forecast in a given week were assigned the median score of all other participants who submitted a forecast that week. The UK crowd forecasting challenge was announced over Twitter and our networks. 
% (e.g. https://www.voice-global.org/public/opportunities/archived/uk-covid-19-crowd-forecasting-challenge/)
In addition, we created a project website, \url{https://crowdforecastr.org}, made weekly posts on Twitter and sent participants who had registered on the online application weekly emails with a reminder and a summary of their past performance. A public leaderboard was available on our website \url{https://epiforecasts.io}. Participants could choose to make a direct forecast as well as an $R_t$ forecast and were counted as two separate forecasters and eligible for prizes twice. Weekly forecasts had to be submitted between Sunday 12pm and Monday 8pm UK time.

@@

Our analysis of the data largely follows the one outlined in \citet{bosseComparingHumanModelbased2022} except for our extension to also consider the natural log of forecasts as a proxy for the exponential growth rate \citet{bosseTransformationForecastsEvaluating2023}. If not otherwise stated, we present results for two-week-ahead forecasts, following the practice adopted by the COVID-19 Forecast Hubs, which found predictive performance to be poor and unreliable beyond this horizon \cite{cramerEvaluationIndividualEnsemble2021, sherrattPredictivePerformanceMultimodel2022a, bracherShorttermForecastingCOVID192021}. We analysed all forecasts stratified by forecast target (cases or deaths), forecast horizon, and forecast approach. We compared the performance of the direct vs. $R_t$ forecasting approach using instances where we had both a direct forecasts and an $R_t$ forecast from the same person.

@@

The median number of unique participants in any given week was 17, the minimum was 6 and the maximum was 51. This was higher than the number of participants in \cite{bosseComparingHumanModelbased2022} (which had a median number of 6, a minimum of 2, and a maximum 10). With respect to the number of submissions from an individual participant, we observed similar patterns as \cite{bosseComparingHumanModelbased2022}: An individual forecaster participated on average in 2.6 weeks out of 13. The median number of submissions from a single individual was one, meaning that similar to \citep{bosseComparingHumanModelbased2022} most forecasters dropped out after their first submission. Only five participants submitted a forecast in ten or more weeks and only two submitted a forecast in all thirteen weeks, one of whom is an author on this study. The number of direct forecasts was higher than the number of $R_t$ forecasts in all weeks (see Figure \ref{fig:num-forecasters}).

\begin{figure*}[ht]
%\centering
\includegraphics[width=0.99\textwidth]{../output/figures/performance.png}
\caption{\bf{Predictive performance across forecast horizons.} A-D: WIS stratified by forecast horizon for cases and deaths on the natural and log scale. E, F: Empirical coverage of the 50\% and 90\% prediction intervals stratified by forecast horizon and target type.}
\label{fig:performance}
\end{figure*}

@@

@@

In terms of the WIS on the natural scale, direct human forecasts performed best at a two-week forecast horizon, while $R_t$ forecasts performed significantly worse than all other approaches (see Table \ref{tab:scores} and Figure \ref{fig:performance}). From June to the end of July, $R_t$ forecasts overpredicted deaths and were noticeable higher than other forecasts, whereas in August, $R_t$ forecasts underpredicted deaths and were substantially lower than other forecasts (see Figure \ref{fig:forecasts-scores}). For four-week-ahead death forecasts, the Hub ensemble performed slightly better than other approaches. 
On the natural scale, the highest scores were concentrated towards the end of the study period, where death incidences were highest, while on the log scale, scores were more evenly distributed across the study period. On the log scale, direct human forecasts, the combined crowd ensemble and the Forecast Hub ensemble showed comparable performance, while $R_t$ forecasts performed again significantly worse.

$R_t$ forecasts show strong underdispersion with empirical coverage of the 50\% prediction intervals being close to zero and smaller than 50\% for the 90\% prediction intervals (see Figure \ref{fig:performance} and Table \ref{tab:scores}). The Forecast Hub ensemble was overdispersed with empirical coverage exceeding nominal coverage, and for the direct human forecasts, empirical coverage was close to nominal coverage.

@@

For cases, where participants could observe the case forecast implied by their $R_t$ forecast, predictive performance was similar between corresponding direct and $R_t$ forecasts for most forecasters who had submitted both (see Figure \ref{fig:comparison-direct-rt-individual}). For deaths, where forecasters could not see the incidence forecast implied by their $R_t$ forecast or manually adjust the case fatality rate, performance of the $R_t$ forecasts was significantly worse than the corresponding direct forecasts for most forecasters. This suggests that the convolution from latent infections implied by the $R_t$ forecasts to deaths did not capture the relationship between cases and deaths accurately in this instance despite performing well in other settings [CITE SOPHIE's paper]. This may be related to the change in the case fatality ratio over the course of the study period due to the rise of the Delta variant starting in May 2021 in the UK which was not accounted for in the simple model used and not available to be altered by forecasters.

@@

In this paper, we presented a follow-up study to \cite{bosseComparingHumanModelbased2022}, analysing human judgement forecasts of cases of and deaths from COVID-19 in the United Kingdom submitted to the European COVID-19 Forecast Hub between the 24th of May and the 16th of August 2021. Human judgement forecasts were generated using two different forecasting approaches, a) direct forecasts of cases and deaths and b) forecasts of the effective reproduction number $R_t$, which were based on output from an open source effective reproduction number estimation model and then used to generate simulated reported cases and deaths using the same model and a simple extension which assumed a static case fatality ration and a delay between cases and reports that was estimated from the data. We found that human judgement forecasts of case numbers, regardless of how they were obtained, showed performance broadly comparable to the ensemble of all forecasts submitted to the European Forecast Hub. For deaths, performance of direct human forecasts was comparable to that of the Forecast Hub ensemble, while $R_t$ forecasts performed significantly worse. Model rankings were not always robust, but sometimes changed depending on how forecasts are evaluated. For example, model rankings could differ when focusing on a different forecast horizon, or analysing WIS values on the natural scale vs. on the log scale (i.e. after log-transforming both forecasts and observations).
